---
title: "Rhododendron Gardens Tree Girth Project"
author: "Michael Dear"
date: "2025-07-20"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
The girths of 30 trees were measured in the Campbell Rhododendron Gardens, Blackheath, NSW. The measurements were made 1.4m from ground level and were recorded to the nearest centimetre. Photographs containing GPS and other ancilliary information were taken of each tree.

## Objectives
The objectives (OB) of this project were to:

1. create an interactive map using Leaflet showing the location of each tree with the girth of each tree displayed as a popup
2. produce a boxplot and histogram of the girth measurements
3. provide a quantitative summary table of the girth measurements.

## References

* https://www.r-bloggers.com/2016/11/extracting-exif-data-from-photos-using-r/ for the basic concept.
* https://youtu.be/TZ8CxNNHiUU?si=qBd84kY_Ep0JwaiS for creating maps using `sf` and `leaflet`.
* https://rstudio.github.io/leaflet/index.html for `leaflet` examples.

# Preparation
## Libraries etc
```{r Libraries etc}
# Clear the environment
rm(list=ls())

library(exiftoolr) # To extract EXIF information from photos; Need to run 'install_exiftoolr()' from R command line after install.packages('exiftoolr')
library(dplyr)
library(sf) # For making geospatial data frames
library(leaflet) # For making interactive maps
library(htmltools) # For map popups
library(ggplot2)
```


## Import and Prepare the Data

Create a list of file names from the photos directory, then read the EXIF data into a data frame.
```{r Import and Prepare the Data}
files <- list.files("./Input/Tree Photos", pattern = "*.JPG")
df_base <- exif_read(paste("./Input/Tree Photos/", files, sep="")) 
```

Merge in the tree girth measurements.
```{r Merge tree girths}
df_base <- read.csv("Input/tree_girths.csv") |> 
                      select(Girth_mm) |>
                      bind_cols(df_base)
```

Select the desired variables and convert to an `sf` object.
```{r create sf object}
df_map <- df_base |> 
  select(FileName, DateTimeOriginal, Girth_mm, GPSLatitude, GPSLongitude) |> 
  st_as_sf(coords = c("GPSLongitude","GPSLatitude"), remove = FALSE, crs = 4326)
```


## OB 1

### Create An Interactive Map

Create a continuous palette function based on  https://rstudio.github.io/leaflet/articles/colors.html with the palette selected from https://colorbrewer2.org/#type=sequential&scheme=RdPu&n=3 .

```{r continuous colour palette}
# Create a continuous palette function
# https://rstudio.github.io/leaflet/articles/colors.html
# https://colorbrewer2.org/#type=sequential&scheme=RdPu&n=3
pal <- colorNumeric(
  palette = "RdPu",
  domain = df_map$Girth_mm)
```

Create a marker size function. This is required since division used within the `radius` parameter of `addCircleMarkers` causes an error.
```{r marker size function}
# Required due to error caused by division when used within 'radius' parameter
size_func <- function(x){x/max(x)*5}
```

Create the map.
```{r Interactive Map}

map <- leaflet(elementId = "Trees", 
             data = df_map, # Same data set will be available to all layers
             options = leafletOptions(minZoom = 15, maxZoom = 20)) |>
  
  # Tiles
  addTiles(group = "OSM (default)") |> # Gives Open Street Map tiles 
  addProviderTiles(providers$CartoDB.Positron, group = "Positron (minimal)") |>
  addProviderTiles(providers$Esri.WorldImagery, group = "World Imagery (satellite)") |>
  setView(150.2864, -33.6268, zoom = 18) |>
  
  # Markers
  addCircleMarkers(popup = ~htmlEscape(paste("Girth: ", Girth_mm, "mm", sep = "")), 
                   # radius = ~sqrt(Girth_mm)*0.1, 
                   radius = ~size_func(Girth_mm), 
                   color = ~pal(Girth_mm),
                   group = "Trees") |>
  
  # Controls
   addLayersControl(
    baseGroups = c(
      "OSM (default)",
      "Positron (minimal)",
      "World Imagery (satellite)"
    ),
    overlayGroups = c("Trees"),
    options = layersControlOptions(collapsed = FALSE)) |>

  # Legend
   addLegend("bottomright", pal = pal, values = ~Girth_mm,
    title = "Girth (mm) at 1.4m",
    labFormat = labelFormat(suffix = "mm"),
    opacity = 1)


map
```



## OB 2

### Boxplot of Girths
```{r Boxplot of Girths}
df_map %>%
  ggplot(aes(y=Girth_mm)) +
  geom_boxplot() +
  scale_x_discrete() +
  labs(title="Boxplot of Tree Girths", y="Girth (mm) at 1.4m")
```


### Histogram of Girths
```{r Boxplot of Girths}
df_map %>%
  ggplot(aes(x=Girth_mm)) +
  geom_histogram(bins=5) +
  labs(title="Histogram of Tree Girths", x="Girth (mm) at 1.4m", y="Frequency")
```



## OB 3

### Summary Table of Girth Statistics
```{r Summary table of Girth statistics}
summary(df_map$Girth_mm)
```

